Le titre

### 2022-09-29 Manuscrit Thèse/Intro et conclusion/Intro.tex
tant de temps ...

As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble. While single NV center are required for most quantum information processes \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble. Some of these applications are pictured in Fig.

The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers.

While most of the NV ensemble experiments work under the assumption that the ensemble is constituted of $N$ independent single NV centers

>>>>

As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble. While single NV center are required for most quantum information processing \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble. Four of these applications are pictured in Fig. [image bio : mohan2010vivo]

The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers \citep{fu2007characterization, mohan2010vivo}. NV centers are good candidates for biomarkers as they emit a bright and stable photoluminesecence, and because diamond nanocrystals present low toxicity \citep{schirhagl2014nitrogen}. Some applications also use the spin degree of freedom the NV centers, either to perform background free fluorescence microscopy \citep{chapman2013background} or to monitor the angular motion of the nanodiamonds \citep{mcguinness2011quantum, feng2021association}. For most of these applications, NV ensemble provide a better visibility than single NV centers and are therefore preferred.

The second NV ensemble application is for quantum sensing \citep{degen2017quantum}. NV centers are predominantly used as magnetic field sensors \citep{rondin2014magnetometry, barry2020sensitivity}, although they can also detect electric fields \citep{dolde2011electric, michl2019robust}, temperature \citep{acosta2010temperature, kucsko2013nanometre} ,crystal strain \citep{ovartchaiyapong2014dynamic, doherty2014electronic} or rotation \citep{ledbetter2012gyroscopes, ajoy2012stable}. Both single and ensemble of NV centers can be used for sensing: single centers offer unmatched spatial resolution \citep{mitchell2020colloquium} down to a few nm \citep{lovchinsky2016nuclear}, while ensemble provide better sensitivities due to the larger number of probes \citep{wolf2015subpicotesla, barry2020sensitivity}.

The third NV ensemble application is nuclear magnetic resonance (NMR) dynamical nuclear polarization \citep{eills2022spin}. Dynamic nuclear polarization consists in increasing the nuclear polarization of a sample by transferring electron polarization to the nuclei bath \citep{abragam1978principles}. Because the NV electronic spin can be optically polarized at room temperature, micro and nano diamonds containing NV centers are promising candidates as hyperpolarization agents \citep{tetienne2021prospects}. Dynamic nuclear polarization of nuclear spins inside the diamond lattice was demonstrated with single NV centers \citep{jacques2009dynamic, smeltzer2009robust} and NV ensemble \citep{king2015room, scheuer2016optically, schwartz2018robust}. Polarization transfer to nuclear spins outside of the diamond is an ongoing challenge \citep{eills2022spin, healey2021polarization}, but dense NV ensemble are believed to be necessary to achieve high external polarization \citep{tetienne2021prospects, rizzato2022polarization}.

Finally, the fourth NV ensemble application is the recent field of spin mechanics \citep{perdriat2021spin}. Similarly to optomechanics \citep{aspelmeyer2014cavity}, the goal of spin mechanics is to couple a quantum system to the mechanical degree of freedom of a macroscopic object. Strong coupling between single NV center spin and external mechanical oscillators has been observed \citep{rabl2009strong, kolkowitz2012coherent}, but coupling of the diamond motion with NV centers spin has so far only been achieved with dense NV ensemble \citep{delord2020spin, pellet2021magnetic, perdriat2022angle}. Coupling several NV centers to the same mechanical oscillator could also lead to interesting magnetic phase transitions \citep{wei2015magnetic, ma2017proposal}.

Most of the NV ensemble experiments discussed previously work under the assumption that the ensemble is constituted of $N$ independent single NV centers. However, as often in physics, \textit{more is different} \citep{anderson1972more}. Interaction between the NV center themselves or with their environment can lead to significantly different behavior between ensemble and single spins \citep{dobrovitski2008decoherence, choi2017observation, kucsko2018critical, giri2018coupled}.

The work presented in this manuscript focuses on some of the
